### 2013-07-30 appa.tex
fml

Fix an expression $\ul r$ of $s$'s and $t$'s with length $n$.  Then for any binary string $\ul b$ of length $n$ we can construct a diagram based on a series of rules.

@@

	Define a pair of red anchors $u, v$, with $u$ to the left of $v$, to be \emph{quasi-connected} if (i) one can draw an arc from $u$ to $v$ contained entirely in one half-map which does not intersect blue lines, and (ii) there is no red anchor $w$ strictly between $u$ and $v$ for which one can draw a similar arc on the same half-map joining $w$ and $v$.

>>>>

Fix an expression $\ul r$ of length $n$.  Then for any binary string $\ul b$ of length $n$ we can construct a diagram based on a series of rules.

@@

	Define a pair of red anchors $u, v$, with $u$ to the left of $v$, to be \emph{quasi-connected} if
	\begin{inparaenum}[(i)]
		\ii one can draw an arc from $u$ to $v$ contained entirely in one half-map which does not intersect blue lines, and
		\ii there is no red anchor $w$ strictly between $u$ and $v$ for which one can draw a similar arc on the same half-map joining $w$ and $v$.
	\end{inparaenum}

### 2013-07-30 paper.tex
fml

The Hecke Algebra (or Iwahori-Hecke Algebra) is an object of significant interest, studied in representation theory, algebraic number theory, and combinatorics.  In 1979, Kazhdan and Lusztig \cite{ref:KaLu1,ref:KaLu2} gave a categorification of the Hecke algebra, using a somewhat complicated construction\footnote{The technique involved using something called \emph{perverse sheaves on the flag variety}.}.  The result became known as the Hecke category.
% If $K = \RR$, the resulting basis of $\HH$ is understood, and called the Kazhdan-Lusztig basis.  However, if $K$ is a field of some other characteristic $p$, the basis is unknown.
Soergel \cite{ref:Soe1,ref:Soe2} simplified this construction significantly in 1990 with an algebraic approach, using the aptly named Soergel bimodules.  The Soergel bimodules permit an easier understanding of this categorification; their advantage is that they are easier to define and are combinatorial in nature.
% Furthermore, Soergel's construction extends the Hecke category to the case where $W$ is an arbitrary Coxeter group, as opposed to a Weyl group.

@@

\begin{definition}
	A connected component which (i) is acyclic, and (ii) does not touch the top or bottom boundaries is called a \emph{barbell}.
\end{definition}
Notice that, by combining homotopy and contraction (see Appendix~\ref{sec:prelim_genrel}), every blue barbell is simply $\barbell{barbell_blue}$.  Similarly, every red barbell is simply $\barbell{barbell_red}$.

@@

Recall that the generators $s$ and $t$ of $W$ act on $R$ by the action given in Appendix~\ref{sec:prelim_background}.  The quantum numbers are relevant to the problem because of the identity
\begin{equation}
	s([n]_x\alpha_s + [n-1]_y\alpha_t) = [n]_x\alpha_s + [n+1]_y\alpha_t
	\label{eq:quantum}
\end{equation}
and its analogous form for $t$.  One can verify that \eqref{eq:quantum} immediately follows from definitions.  This gives us a robust way to express the results of action on $\alpha_s$ or $\alpha_t$ by the elements of $W$.  For example, $ststst(\alpha_s) = ststst([1]_x\alpha_s + [0]_y\alpha_t) = [7]_x\alpha_s + [6]_y\alpha_t$.

Now \eqref{eq:quantum} motivates us to make the following definition, which we use in Section~\ref{sec:res_alt_compute}.
\begin{definition}
	For any reduced expression $\ul r$, we define the quantum polynomial $Q_{\ul r}$ by:
	\[
		Q_{\AA_n} = 
		\begin{cases}
			[n+1]_x \alpha_s + [n]_y \alpha_t & \text{if $n$ is odd} \\
			[n]_x \alpha_s + [n+1]_y \alpha_t & \text{if $n$ is even}
		\end{cases}
	\]
	when $\ul r = \AA_n$, and otherwise,
	\[
		Q_{\AA_n'} =
		\begin{cases}
			[n]_x \alpha_s + [n+1]_y \alpha_t & \text{if $n$ is even} \\
			[n+1]_x \alpha_s + [n]_y \alpha_t & \text{if $n$ is odd}.
		\end{cases}
	\]
\end{definition}

@@

\section{Problem Statement}
\label{sec:probstate}
\begin{definition}
	A map is \emph{idempotent} if multiplying it with itself returns the original map.  Two idempotents are \emph{orthogonal} if their product is zero.
\end{definition}
If we can find two half-maps $A$ and $B$ (with the same top and base) for which $AB$ is the identity, then $(BA)^2=BABA=BA$ is an idempotent.  More generally, however, if $AB = c$ for some scalar $c \neq 0$, then \[ \left( \frac{BA}{c} \right)\left( \frac{BA}{c} \right) = \frac{BABA}{c^2} = \frac{BA}{c} \] is also an idempotent.  In other words, any two maps with nonzero product can be used to construct an idempotent.

However, if we are interested in finding sets of \emph{pairwise orthogonal} idempotents, then this becomes a linear algebra problem in which we require a matrix of all possible products.  The construction of such a set of idempotents is closely related to the decomposition of these bimodules, which in turn gives information about the Soergel bimodules\footnote{Actually, the Soergel bimodules may be defined as the images of idempotents in the endomorphism rings of the Bott-Samelson bimodules.}.

@@

The proof proceeds in two steps.  The first is to resolve the case $\ell = n$.  We do so by using induction on $n$, and applying \eqref{eq:top}.  By splitting into the case where $\ul b[n] = 1$ and $\ul b[n] = 0$, we get two binomial coefficients; the computation is to verify that by Pascal's identity they sum to the desired coefficient.
The second step is to reduce the case where $\ell \neq n$ to the previous step using Proposition~\ref{prop:s_top}.  We give the proof in Appendix~\ref{sec:count_map_computation}, on page~\pageref{sec:count_map_computation}.

@@

Below we present three results which show that the possible outputs of a map with base $\AA_n$ are structurally very limited.
\begin{restatable}[Bubble Lemma]{lemma}{bubble}
	Consider a bubble in a full-diagram with base $\AA_n$.  The bubble holds exactly one caterpillar, whose color is opposite that of the bubble.
	\label{thm:bubble}
\end{restatable}
\begin{restatable}[Caterpillar Lemma]{lemma}{caterpillar}
	Any caterpillar in a full-diagram with base $\AA_n$ evaluates in the form $x^my^n\alpha_s$ if it is blue, or $x^my^n\alpha_t$ if it is red.
	\label{thm:caterpillar}
\end{restatable}
\begin{restatable}[Pasture Lemma]{lemma}{pasture}
	Consider a full-diagram with $N+1$ pastures (numbered $0$, $1$, \dots, $N$) with base $\AA_n$ and top $\ul w$.  Suppose further that $N \ge 1$ (or equivalently, $\ul w \neq \varnothing$).  Then
	\begin{enumerate}[(i)]
		\ii Pastures $1$ through $N-1$ are either empty or contain an attached component.
		\ii Pasture $N$ is empty if $\ul w[N] = \AA_n[n]$, and contains a single caterpillar otherwise.  The caterpillar is blue if $\AA_n[n]$ is $s$, and red otherwise.
	\end{enumerate}
	\label{thm:pasture}
\end{restatable}

@@

The Bubble Lemma is the key combinatorial observation, which hinges on the fact that any two anchors with the same color have a third anchor of the other color between them.  The other two lemmas follow directly from it.
The details of the proofs are given in Appendix~\ref{sec:lemmata_proof}, on page~\pageref{sec:lemmata_proof}.

@@

Theorem~\ref{thm:alt_struct} motivates the following definition, which is relevant for Proposition~\ref{thm:recurse}.
\begin{definition}
	If a nonzero map $M$ with base $\AA_n$ is presented as in Theorem~\ref{thm:alt_struct}, then we call the term $x^{m_x}y^{m_y}\alpha_s^{m_s}\alpha_t^{m_t}$ the \emph{fluff} of $M$, and call the polynomial (either $1$ or $Q_{\ul w}$) the \emph{face} of $M$. 
\end{definition}
\begin{remark*}
	According to Theorem~\ref{thm:alt_struct}, a nonzero map $\MM_{\AA_n}$ has face $1$ and only if its top has the same last character as $\AA_n$.
\end{remark*}

@@

What about the original problem of arbitrary $\ul r$?  It was originally hoped that, by understanding $\AA_n$ and $\SS_n$, one could derive the general formula (perhaps by viewing strings of $s$'s in $\ul r$ as a single $s$ with a formula for $\SS_n$, and then aggregating all this information with a formula for $\AA_n$).  Unfortunately, the lemmata of Section~\ref{sec:res_alt_restrict} show that in fact, maps with base $\AA_n$ are quite restricted.  So, either the results of Section~\ref{sec:res_alt_compute} will have to be generalized significantly (in the context of maps outside those of base $\AA_n$), or an entirely new approach will be required.

>>>>

The Hecke Algebra (or Iwahori-Hecke Algebra) is an object of significant interest, studied in representation theory, algebraic number theory, and combinatorics.  In 1979, Kazhdan and Lusztig \cite{ref:KaLu1,ref:KaLu2} gave a categorification of the Hecke algebra, using a somewhat complicated construction\footnote{The technique involved using something called \emph{perverse sheaves on the flag variety}.}.  The result became known as the Hecke category.
Soergel \cite{ref:Soe1,ref:Soe2} simplified this construction significantly in 1990 with an algebraic approach, using the aptly named Soergel bimodules.  The Soergel bimodules permit an easier understanding of this categorification; their advantage is that they are easier to define and are combinatorial in nature.
% Furthermore, Soergel's construction extends the Hecke category to the case where $W$ is an arbitrary Coxeter group, as opposed to a Weyl group.

@@

\begin{definition}
	A connected component which
	\begin{inparaenum}[(i)]
		\ii is acyclic, and 
		\ii does not touch the top or bottom boundaries
	\end{inparaenum} is called a \emph{barbell}.
\end{definition}
Notice that, by combining homotopy and contraction (see Appendix~\ref{sec:prelim_genrel}), every blue barbell is simply $\barbell{barbell_blue}$.  Similarly, every red barbell is simply $\barbell{barbell_red}$.

@@

Recall that the generators $s$ and $t$ of $W$ act on $R$ by the action given in Appendix~\ref{sec:prelim_background}.  The quantum numbers are relevant to the problem because of the identity
\begin{equation}
	s([n-1]_x\alpha_s + [n]_y\alpha_t) = [n]_x\alpha_s + [n+1]_y\alpha_t
	\label{eq:quantum}
\end{equation}
and its analogous form for $t$.  One can verify that \eqref{eq:quantum} immediately follows from definitions.  This gives us a robust way to express the results of action on $\alpha_s$ or $\alpha_t$ by the elements of $W$.  For example, $ststst(\alpha_s) = ststst([1]_x\alpha_s + [0]_y\alpha_t) = [7]_x\alpha_s + [6]_y\alpha_t$.

Now \eqref{eq:quantum} motivates us to make the following definition, which we use in Section~\ref{sec:res_alt_compute}.
\begin{definition}
	For any reduced expression $\ul r$, we define the quantum polynomial $Q_{\ul r}$ by $Q_{\AA_n} = [n+1]_x \alpha_s + [n]_y \alpha_t$ when $\ul r = \AA_n$, and otherwise, $Q_{\AA_n'} = [n]_x \alpha_s + [n+1]_y \alpha_t$.
\end{definition}

@@

\section{Problem Statement}
\label{sec:probstate}
\begin{definition}
	A map is \emph{idempotent} if multiplying it with itself returns the original map.  Two idempotents are \emph{orthogonal} if their product is zero.
\end{definition}
If we can find two half-maps $A$ and $B$ for which $AB$ is the identity, then $(BA)^2=BABA=BA$ is an idempotent.  More generally, however, if $AB = c$ for some scalar $c \neq 0$, then \[ \left( \frac{BA}{c} \right)\left( \frac{BA}{c} \right) = \frac{BABA}{c^2} = \frac{BA}{c} \] is also an idempotent.  In other words, any two maps with nonzero product can be used to construct an idempotent.

However, if we are interested in finding sets of \emph{pairwise orthogonal} idempotents, then this becomes a linear algebra problem in which we require a matrix of all products.  The construction of such a set of idempotents is closely related to the decomposition of these bimodules, which in turn gives information about the Soergel bimodules\footnote{Actually, the Soergel bimodules may be defined as the images of idempotents in the endomorphism rings of the Bott-Samelson bimodules.}.

@@

The proof proceeds in two steps.  The first is to resolve the case $\ell = n$.  We do so by using induction on $n$, and applying \eqref{eq:top}.  By splitting into the case where $\ul b[n] = 1$ and $\ul b[n] = 0$, we get two binomial coefficients; the computation is to verify that we can then apply Pascal's identity.
The second step is to reduce the case where $\ell \neq n$ to the previous step using Proposition~\ref{prop:s_top}.  We give the proof in Appendix~\ref{sec:count_map_computation}, on page~\pageref{sec:count_map_computation}.

@@

Below we present three results which show that the possible outputs of a map with base $\AA_n$ all take certain forms.
\begin{restatable}[Bubble Lemma]{lemma}{bubble}
	Consider a bubble in a full-diagram with base $\AA_n$.  The bubble holds exactly one caterpillar, whose color is opposite that of the bubble.
	\label{thm:bubble}
\end{restatable}
\begin{restatable}[Caterpillar Lemma]{lemma}{caterpillar}
	Any caterpillar in a full-diagram with base $\AA_n$ evaluates in the form $x^my^n\alpha_s$ if it is blue, or $x^my^n\alpha_t$ if it is red.
	\label{thm:caterpillar}
\end{restatable}
\begin{restatable}[Pasture Lemma]{lemma}{pasture}
	Consider a full-diagram with $N+1$ pastures (numbered $0$, $1$, \dots, $N$) with base $\AA_n$ and top $\ul w$.  Suppose further that $N \ge 1$ (or equivalently, $\ul w \neq \varnothing$).  Then
	\begin{enumerate}[(i)]
		\ii Pastures $1$ through $N-1$ are either empty or contain an attached component.
		\ii Pasture $N$ is empty if $\ul w[N] = \AA_n[n]$, and contains a single caterpillar otherwise.  The caterpillar is blue if $\AA_n[n]$ is $s$, and red otherwise.
	\end{enumerate}
	\label{thm:pasture}
\end{restatable}

@@

The Bubble Lemma is the key combinatorial observation, which hinges on the fact that any two anchors with the same color have a third anchor of the other color between them.  The other two lemmas follow directly from it.
The proofs are given in Appendix~\ref{sec:lemmata_proof}, on page~\pageref{sec:lemmata_proof}.

@@

Theorem~\ref{thm:alt_struct} motivates the following definition, which is relevant for Proposition~\ref{thm:recurse}.
\begin{definition}
	If a nonzero map $M$ with base $\AA_n$ is presented as in Theorem~\ref{thm:alt_struct}, then we call the term $x^{m_x}y^{m_y}\alpha_s^{m_s}\alpha_t^{m_t}$ the \emph{fluff} of $M$, and call the polynomial (either $1$ or $Q_{\ul w}$) the \emph{face} of $M$. 
\end{definition}
\begin{remark*}
	According to Theorem~\ref{thm:alt_struct}, a nonzero map $\MM_{\AA_n}$ has face $1$ if and only if its top has the same last character as $\AA_n$.
\end{remark*}

@@

What about the original problem of arbitrary $\ul r$?  It was originally hoped that, by understanding $\AA_n$ and $\SS_n$, one could derive the general formula (perhaps by viewing strings of $s$'s in $\ul r$ as a single $s$ with a formula for $\SS_n$, and then aggregating all this information with a formula for $\AA_n$).  Unfortunately, the lemmata of Section~\ref{sec:res_alt_restrict} show that in fact, maps with base $\AA_n$ have very special properties.  So, either the results of Section~\ref{sec:res_alt_compute} will have to be generalized significantly (in the context of maps outside those with base $\AA_n$), or an entirely new approach will be required.
